Pillar 1 - Sequential Pattern Mining

### 2019-01-28 Pillar 1 - Sequential Pattern Mining.Rmd
Creating Sequences

# Association Rules
For a first exploration of arules, we conceptualise our framework like this:
transaction = student
item = course

@@

```{r AR data prep 1}
d_transactions <- d_course %>%

@@

  # Join with transcripts.
  inner_join(d_transcript, by = "Course ID")
```

```{r AR data prep 2}
# Create transactions (student level)
transactions <- d_transactions %>%
  group_by(`Student ID`) %>%
    summarize(list_course = list(`Course ID`)) %>%
  ungroup

transactions <- as(transactions$list_course, "transactions")
```

```{r AR data prep 3 (PF)}
d_transactions_PF <- d_transactions %>%

  # Take outcome of course into account (pass or fail)
  mutate(PF = case_when(Grade <  5.5 ~ "fail",
                        Grade >= 5.5 ~ "pass"),
         `Course ID` = paste(`Course ID`, PF, sep = "_"))

  # Create transactions with pass/fail
transactions_PF <- d_transactions_PF %>%
  group_by(`Student ID`) %>%
    summarize(list_course = list(`Course ID`)) %>%
  ungroup

transactions_PF <- as(transactions_PF$list_course, "transactions")
```

@@

#Mining-sequences:
The parameter values are: 
support:  a numeric value specifying the minimum support of a sequence (default 0.1, range [0,1]).
maxsize:  an integer value specifying the maximum number of items of an element of a sequence (default 10, range > 0).
maxlen:   an integer value specifying the maximum number of elements of a sequence (default 10, range > 0).
mingap:   an integer value specifying the minimum time difference between consecutive elements of a sequence (default none, range >= 1).
maxgap:   an integer value specifying the maximum time difference between consecutive elements of a sequence (default none, range >= 0).
maxwin:   an integer value specifying the maximum time difference between any two elements of a sequence (default none, range >= 0).

##Preparing data frame:
```{r}
fail_threshold <- 6.5

category <-quo(Low_grade) #Outcome or Low_grade or Discrete_grade or #None

#this is new-----
transcript_to_use <- quo(distributed_transcript) #d_transcript or distributed_transcript 
#----------------
d_sequence_transactions <- d_course %>%

  left_join(d_transcript, by = "Course ID") %>%
  # filter(!is.na(Grade)) %>% #remove NaN
  # filter(Grade != -1) %>% #remove NG
  mutate(
    Outcome = 
           case_when( Grade == -1   ~"Fail", #if we don't do this NG satisfies the condition for pass v
                      Grade  > 5.5 ~"Pass",
                               T~"Fail"),
    Low_grade =
           case_when(Grade == -1 ~"low score", #if we don't do this NG satisfies the condition for pass v
                      Grade > fail_threshold ~"high score",
                               T~"low score"),
    Discrete_grade =
           case_when(Grade <= 5.4             ~ 5.0,
                     between(Grade, 5.5, 6.4) ~ 6.0,
                     between(Grade, 6.5, 7.4) ~ 7.0,
                     between(Grade, 7.5, 8.4) ~ 8.0,
                     between(Grade, 8.5, 9.4) ~ 9.0,
                     Grade >= 9.5             ~ 10.0,
                     T ~ -99999    #this is a control, we should not have any grades here
                     ),
    None= "" # this might not be optimal
         ) %>%

#checking results
as(transactions, "data.frame")
```

##Rule induction
##Creating look-up table
```{r}
lookup_table <- d_course %>%

>>>>

# Association Rules and Sequence Rules
For a first exploration of arules, we conceptualise our framework like this:
transaction = student
item = course

@@

```{r AR data prep 1}

#
# Threshold: low vs high grade
low_grade <- 6.5

#
# Transactions
d_transactions <- d_course %>%

@@

  # Join with transcripts.
  select(- Period) %>%
  inner_join(d_transcript, by = "Course ID") %>%

  # Identifying sequences
  rename(sequenceID = `Student ID`) %>%

  # Identifying time of event (sequence)
  mutate(
    Period  = substr(Period, 1, 1),
    eventID = as.numeric(paste(Year_numerical, Period, sep = ""))
    ) %>%

  # Identifying item
  mutate(
    PF = case_when( Grade <  5.5 ~ "fail",
                    Grade >= 5.5 ~ "pass"),
    HL = case_when( Grade <  low_grade ~ "low",
                    Grade >= low_grade ~ "high"),

    item_PF = paste(`Course ID`, PF, sep = "_"),
    item_HL = paste(`Course ID`, HL, sep = "_")
    ) %>%
  rename(
    item = `Course ID`
  )
```

### Transactions for Apriori
```{r AR data prep 2}
transactions <- d_transactions %>%
  group_by(
    sequenceID
    ) %>%
  summarize(
    list_item = list(item)
    ) %>%
  ungroup

transactions <- as(transactions$list_item, "transactions")
```

```{r AR data prep 3 (PF)}
transactions_PF <- d_transactions %>%
  group_by(
    sequenceID
    ) %>%
  summarize(
    list_item = list(item_PF)
    ) %>%
  ungroup

transactions_PF <- as(transactions_PF$list_item, "transactions")
```

```{r AR data prep 4 (HL)}
transactions_HL <- d_transactions %>%
  group_by(
    sequenceID
    ) %>%
  summarize(
    list_item = list(item_HL)
    ) %>%
  ungroup

transactions_HL <- as(transactions_HL$list_item, "transactions")
```

### Sequences for CSPADE
```{r}
sequences_temp <- d_transactions %>%
  group_by(
    sequenceID,
    eventID
    ) %>%
  summarize(
    list_item = list(item)
    ) %>%
  arrange(
    sequenceID,
    eventID
    )

sequences             <- as(sequences_temp$list_item, "transactions")
sequences@itemsetInfo <- as.data.frame(select(sequences_temp, sequenceID, eventID))

#checking results
#inspect(head(sequences, 10))
```

```{r}
sequences_temp <- d_transactions %>%
  group_by(
    sequenceID,
    eventID
    ) %>%
  summarize(
    list_item = list(item_PF)
    ) %>%
  arrange(
    sequenceID,
    eventID
    )

sequences_PF             <- as(sequences_temp$list_item, "transactions")
sequences_PF@itemsetInfo <- as.data.frame(select(sequences_temp, sequenceID, eventID))

#checking results
#inspect(head(sequences_PF, 10))
```

```{r}
sequences_temp <- d_transactions %>%
  group_by(
    sequenceID,
    eventID
    ) %>%
  summarize(
    list_item = list(item_HL)
    ) %>%
  arrange(
    sequenceID,
    eventID
    )

sequences_HL             <- as(sequences_temp$list_item, "transactions")
sequences_HL@itemsetInfo <- as.data.frame(select(sequences_temp, sequenceID, eventID))

#checking results
#inspect(head(sequences_HL, 10))
```

@@

## CSPADE Algorithm

The parameter values are: 
support:  a numeric value specifying the minimum support of a sequence (default 0.1, range [0,1]).
maxsize:  an integer value specifying the maximum number of items of an element of a sequence (default 10, range > 0).
maxlen:   an integer value specifying the maximum number of elements of a sequence (default 10, range > 0).
mingap:   an integer value specifying the minimum time difference between consecutive elements of a sequence (default none, range >= 1).
maxgap:   an integer value specifying the maximum time difference between consecutive elements of a sequence (default none, range >= 0).
maxwin:   an integer value specifying the maximum time difference between any two elements of a sequence (default none, range >= 0).

##Creating look-up table
```{r}
lookup_table <- d_course %>%
